---
title: "Bond Sensitivity"
author: "Po-Sheng Lee"
date: "2020/3/11"
output: html_document

---

```{r markdown setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

## The latex syntax will not show in the MD file on github, only fully rendered html with MathJax will be shown

```

```{r library}
library(quantmod)
library(tidyverse)
library(magrittr)
library(MTS)
library(purrr)
library(tinytex)
```

This post is intended to explore how they change in response to change in discount rate. First we will lay out how bond is valued and how it react to small **change in discount rate (YTM)**.

### Bond Valuation and Sensitivity to Discout Rate

Present value of bond can be estimated by

$$PV = \frac{PMT}{r}(1- \frac{1}{(1+r)^{T}}) + \frac{Face Value}{(1+r)^{T}}$$

and the sensitivity of bond value to change in discount rate can be derived from partial derivative of PV, which is 

$$\frac{\partial PV}{\partial r} = PMT(- \frac{1}{r^{2}} + \frac{1}{r^{2}(1+r)^{T}} + \frac{T}{r(1+r)^{T+1}}) - \frac{FV \times T}{(1+r)^{T+1}}$$ 

and the percentage of change can be written as 

$$\frac{(\frac{\partial PV}{\partial r})}{PV} = f(r,T,PMT,FV) = \frac{PMT(- \frac{1}{r^{2}} + \frac{1}{r^{2}(1+r)^{T}} + \frac{T}{r(1+r)^{T+1}}) - \frac{FV \times T}{(1+r)^{T+1}}}{\frac{PMT}{r}(1- \frac{1}{(1+r)^{T}}) + \frac{Face Value}{(1+r)^{T}}}$$ 

###Function to Calculate 

The function to calculate present value of bond is given in the `R code` below:

```{r Present Value of the bond, echo=TRUE}

bondprice <- function(r, n, c = 0, fv = 0){
  ## r is discount rate or yield to maturity
  ## n is period of payment to maturity 
  ## c is coupon payment
  ## fv is face value
  pv <- c/r*(1-1/(1+r)^n) + fv/(1+r)^n
  return(pv)
}

```

The function to calculate $f(r,T,PMT,FV)$ is written as the `R code` below:

```{r bond sensitivity, echo=TRUE}

bondsens <- function(r, n, c = 0, fv = 0) {
  ## r is discount rate or yield to maturity
  ## n is period of payment to maturity 
  ## c is coupon payment
  ## fv is face value
  pv <- c/r*(1-1/(1+r)^n) + fv/(1+r)^n
  dpvdr <- c*(-r^-2 + r^-2*(1+r)^-n + n*r^-1*(1+r)^(-n-1)) - n*fv*(1+r)^(-n-1)
  sensrate <- dpvdr/pv
  return(sensrate)
}


```

###Plotting Bond Price

####Bond price of different coupon at different discount rate

The following plot illustrates how bond price reacts to different interest rate.

```{r bond price diff coupon}

r_list <- seq(0,1, 0.01)

Bprice_coupon4 <- map_dbl(r_list, bondprice, n = 10, c = 40, fv = 1000)
Bprice_coupon8 <- map_dbl(r_list, bondprice, n = 10, c = 80, fv = 1000)
Bprice_coupon12 <- map_dbl(r_list, bondprice, n = 10, c = 120, fv = 1000)

r_Bprice_diffcoupon <- as_data_frame(cbind(Bprice_coupon4, Bprice_coupon8, Bprice_coupon12))

r_Bprice_diffcoupon %>%
  ggplot(aes(x = r_list)) +
  geom_point(aes(y = Bprice_coupon4), color = "red") + 
  geom_point(aes(y = Bprice_coupon8), color = "yellow") +
  geom_point(aes(y = Bprice_coupon12), color = "green") +
  labs( x = "interest rate", y = "Bond price", 
        title = "Different Coupon Rate", subtitle = "R = 4%, Y = 8%, G = 12%")

```

####Bond price of zero coupon with different time to maturity

```{r bond price diff maturity}

r_list <- seq(0,1, 0.01)

Bprice_period5 <- map_dbl(r_list, bondprice, n = 5, c = 0, fv = 1000)
Bprice_period10 <- map_dbl(r_list, bondprice, n = 10, c = 0, fv = 1000)
Bprice_period15 <- map_dbl(r_list, bondprice, n = 15, c = 0, fv = 1000)

r_Bprice_diffperiod <- as_data_frame(cbind(Bprice_period5, Bprice_period10, Bprice_period15))

r_Bprice_diffperiod %>%
  ggplot(aes(x = r_list)) +
  geom_point(aes(y = Bprice_period5), color = "red") + 
  geom_point(aes(y = Bprice_period10), color = "yellow") +
  geom_point(aes(y = Bprice_period15), color = "green") +
  labs( x = "interest rate", y = "Bond price", 
        title = "Different Time to Maturity", subtitle = "R: 5 years, Y: 10 years, G = 15 years")

```
####Bond price of different time to maturity with coupon

```{r bond price}

r_list <- seq(0,1, 0.01)

Bprice_period5wc <- map_dbl(r_list, bondprice, n = 5, c = 80, fv = 1000)
Bprice_period10wc <- map_dbl(r_list, bondprice, n = 10, c = 80, fv = 1000)
Bprice_period15wc <- map_dbl(r_list, bondprice, n = 15, c = 80, fv = 1000)

r_Bprice_diffperiodwc <- as_data_frame(cbind(Bprice_period5wc, Bprice_period10wc, Bprice_period15wc))

r_Bprice_diffperiodwc %>%
  ggplot(aes(x = r_list)) +
  geom_point(aes(y = Bprice_period5wc), color = "red") + 
  geom_point(aes(y = Bprice_period10wc), color = "yellow") +
  geom_point(aes(y = Bprice_period15wc), color = "green") +
  labs( x = "interest rate", y = "Bond price", 
        title = "Different Time to Maturity", subtitle = "R: 5 years, Y: 10 years, G = 15 years")

```


###Plotting Bond Price Sensitivity  

The following plots are bond price sensitivity from discount rate 0 to 100% (presented as 0 to 1 on the x-axis). Different colors means bonds with different characteristics, yet all face value are assumed to be 1000.

####Price Sensitivity of different coupon rate

The following plot is the price sensitivity of bonds with different coupon rates. The red points denote 4%; yellow is 8% and green is 12%. I assume annual payment for 10 years, which means there is 10 payment and coupon payment is 40, 80 and 120 respectively.

Price sensitivity is a negative number, which means as discount rate goes up, the bond price goes down. And the changes is more drastic in a low discount rate environment. These features are shared across all bonds. 

In this plot, what we see is bond with lower coupon rate react more aggresively to the change of discount rate. This is essentially because bond with lower coupon rate rely more on the final payment (face value) in the valuation.

```{r plot different coupon rate}

Bsens_coupon4 <- map_dbl(r_list, bondsens, n = 10, c = 40, fv = 1000)
Bsens_coupon8 <- map_dbl(r_list, bondsens, n = 10, c = 80, fv = 1000)
Bsens_coupon12 <- map_dbl(r_list, bondsens, n = 10, c = 120, fv = 1000)

r_diffcoupon <- as_data_frame(cbind(r_list, Bsens_coupon4, Bsens_coupon8, Bsens_coupon12))

r_diffcoupon %>%
  ggplot(aes(x = r_list)) +
  geom_point(aes(y = Bsens_coupon4), color = "red") + 
  geom_point(aes(y = Bsens_coupon8), color = "yellow") +
  geom_point(aes(y = Bsens_coupon12), color = "green") +
  labs( x = "interest rate", y = "derivitive of Bond price / Bond Price", 
        title = "Different Coupon Rate", subtitle = "R = 4%, Y = 8%, G = 12%")
```

####Price Sensitivity of different time to maturity

Longer time to maturity means one have to wait more time to realize the face value and therefore the value is discounted more times before one gets the final payment. If we are dealing with zero coupon bond, it is easy to deduct that bond with longer time to maturity would react more aggresively to the change in discount rate.

The red is 5 years to maturity; yellow is 10 years and green is 15 years.

```{r plot zero coupon different payment period}
Bsens_period5noC <- map_dbl(r_list, bondsens, n = 5, c = 0, fv = 1000)
Bsens_period10noC <- map_dbl(r_list, bondsens, n = 10, c = 0, fv = 1000)
Bsens_period15noC <- map_dbl(r_list, bondsens, n = 15, c = 0, fv = 1000)

r_diffperiod_no_coupon <- as_data_frame(cbind(r_list, Bsens_period5noC, Bsens_period10noC, Bsens_period15noC))

r_diffperiod_no_coupon %>%
  ggplot(aes(x = r_list)) +
  geom_point(aes(y = Bsens_period5noC), color = "red") + 
  geom_point(aes(y = Bsens_period10noC), color = "yellow") +
  geom_point(aes(y = Bsens_period15noC), color = "green") +
  labs( x = "interest rate", y = "derivitive of Bond price / Bond Price",
        title = "Different Maturity w/o coupon payment", subtitle = "R = 5, Y = 10, G = 15")

```

When comparing coupon payment bond with different time to maturity, things are more complicated. To start with, longer time to maturity means more coupon to be paid, which will increase bond price while longer time to maturity also means the final payment get longer to realized, which will decrease bond price. 

As for price sensitivity, we can follow the same logic. The final payment will constitute a larger part of bond price when the time to maturity is shorter, making it more susceptible to discount rate change. Nonetheless, when the time to maturity is shorter, it also means that final payment get discounted less times, hence making it react less aggresive to change in discount rate.

The result is a plot with multiple intersection as shown below.

```{r plot different payment period with coupon payment}
Bsens_period5 <- map_dbl(r_list, bondsens, n = 5, c = 40, fv = 1000)
Bsens_period10 <- map_dbl(r_list, bondsens, n = 10, c = 40, fv = 1000)
Bsens_period20 <- map_dbl(r_list, bondsens, n = 15, c = 40, fv = 1000)

r_diffperiod <- as_data_frame(cbind(r_list, Bsens_period5, Bsens_period10, Bsens_period20))

r_diffperiod %>%
  ggplot(aes(x = r_list)) +
  geom_point(aes(y = Bsens_period5), color = "red") + 
  geom_point(aes(y = Bsens_period10), color = "yellow") +
  geom_point(aes(y = Bsens_period20), color = "green") +
  labs( x = "interest rate", y = "derivitive of Bond price / Bond Price",
        title = "Different Maturity w/coupon payment", subtitle = "R = 5, Y = 10, G = 20")
```




