---
title: "Bond Sensitivity"
output: html_notebook
---

This is a notebook for the corporate finacen class. 

```{r, echo=FALSE, message=FALSE}
library(quantmod)
library(tidyverse)
library(magrittr)
library(MTS)
library(purrr)
```

Given that present value of bond is estimated by $PV = \frac{PMT}{r}(1- \frac{1}{(1+r)^{T}}) + \frac{Face Value}{(1+r)^{T}} $ and the sensitivity of bond value to change in discount rate can be derived from partial derivative of PV, which is $ \frac{\partial PV}{\partial r} = PMT(- \frac{1}{r^{2}} + \frac{1}{r^{2}(1+r)^{T}} - \frac{T}{r(1+r)^{T+1}}) - \frac{FV \times T}{(1+r)^{T+1}} $ and the percentage of change can be written as $ \frac{(\frac{\partial PV}{\partial r})}{PV} = f(r,T,PMT,FV) = \frac{PMT(- \frac{1}{r^{2}} + \frac{1}{r^{2}(1+r)^{T}} - \frac{T}{r(1+r)^{T+1}}) - \frac{FV \times T}{(1+r)^{T+1}}}{\frac{PMT}{r}(1- \frac{1}{(1+r)^{T}}) + \frac{Face Value}{(1+r)^{T}}} $ constructing function and plot different plots.


```{r bond sensitivity}

bondsens <- function(r, n, c = 0, fv = 0) {
  ## r is discount rate or yield to maturity
  ## n is period of payment to maturity 
  ## c is coupon payment
  ## fv is face value
  pv <- c/r*(1-1/(1+r)^n) + fv/(1+r)^n
  dpvdr <- c*(-r^-2 + r^-2*(1+r)^-n + n*r^-1*(1+r)^(-n-1)) - n*fv*(1+r)^(-n-1)
  sensrate <- dpvdr/pv
  return(sensrate)
}


```

```{r bond bondsens plot}

r_list <- seq(0,1, 0.01)
Bsens_coupon4 <- map_dbl(r_list, bondsens, n = 20, c = 20, fv = 1000)
Bsens_coupon8 <- map_dbl(r_list, bondsens, n = 20, c = 40, fv = 1000)
Bsens_coupon12 <- map_dbl(r_list, bondsens, n = 20, c = 60, fv = 1000)

r_diffcoupon <- as_data_frame(cbind(r_list, Bsens_coupon4, Bsens_coupon8, Bsens_coupon12))

r_diffcoupon %>%
  ggplot(aes(x = r_list)) +
  geom_point(aes(y = Bsens_coupon4), color = "red") + 
  geom_point(aes(y = Bsens_coupon8), color = "yellow") +
  geom_point(aes(y = Bsens_coupon12), color = "green") +
  labs( x = "interest rate", y = "derivitive of Bond price / Bond Price", 
        title = "Different Coupon Rate", subtitle = "R = 4, Y = 8, G = 12")

Bsens_period5 <- map_dbl(r_list, bondsens, n = 10, c = 40, fv = 1000)
Bsens_period10 <- map_dbl(r_list, bondsens, n = 20, c = 40, fv = 1000)
Bsens_period20 <- map_dbl(r_list, bondsens, n = 40, c = 40, fv = 1000)

r_diffperiod <- as_data_frame(cbind(r_list, Bsens_period5, Bsens_period10, Bsens_period20))

r_diffperiod %>%
  ggplot(aes(x = r_list)) +
  geom_point(aes(y = Bsens_period5), color = "red") + 
  geom_point(aes(y = Bsens_period10), color = "yellow") +
  geom_point(aes(y = Bsens_period20), color = "green") +
  labs( x = "interest rate", y = "derivitive of Bond price / Bond Price",
        title = "Different Maturity w/coupon payment", subtitle = "R = 5, Y = 10, G = 20")

Bsens_period5noC <- map_dbl(r_list, bondsens, n = 10, c = 0, fv = 1000)
Bsens_period10noC <- map_dbl(r_list, bondsens, n = 20, c = 0, fv = 1000)
Bsens_period20noC <- map_dbl(r_list, bondsens, n = 40, c = 0, fv = 1000)

r_diffperiod_no_coupon <- as_data_frame(cbind(r_list, Bsens_period5noC, Bsens_period10noC, Bsens_period20noC))

r_diffperiod_no_coupon %>%
  ggplot(aes(x = r_list)) +
  geom_point(aes(y = Bsens_period5noC), color = "red") + 
  geom_point(aes(y = Bsens_period10noC), color = "yellow") +
  geom_point(aes(y = Bsens_period20noC), color = "green") +
  labs( x = "interest rate", y = "derivitive of Bond price / Bond Price",
        title = "Different Maturity w/o coupon payment", subtitle = "R = 5, Y = 10, G = 20")

```


